Abstract

### 2018-05-25 ack.tex
thesis done

I have received an incredible amount of support during my time as an undergraduate student. 
First and foremost, I'd like to thank my advisor, Dr. Sergey Fomel, for his helpful guidance over the past couple of years. 
I genuinely believe that working with him has beneficially shaped the trajectory of my life.
Through his guidance and the opportunities he has given me, he has helped me to realize how much I truly enjoy learning, showed me the importance of reproducible computational research, introduced me new ways of thinking about and approaching problems, and inspired me to continue my studies in graduate school.
For all that, I will be forever grateful.

@@

I would especially like to thank the members of the Texas Consortium for Computational Seismology (TCCS), of which I have been a part for the past two years.
%Thank you for all your guidance, and for being there for the best memories during my undergraduate career.
%Thank you for your intellectually stimulating conversations, life advice, and making me feel as though I am ``part of the group'', despite the fact that I was not a graduate student.
Thank you all for your intellectually stimulating conversations, life advice, and for including me as ``part of the group'', despite the fact that I was only an undergraduate student.
%You are an incredibly enjoyable and intellectually stimulating group, and I am glad I have had the opportunity to cross paths with all of you in life.
I will forever cherish the friendships and memories I have made with all of you during my time as a part of TCCS. %Many of the best memories of my time at UT Austin are from being a part of TCCS, %come from the %time I have spent as a part of TCCS.
%You are an incredibly ingenuitive and enjoyable group, and I am fortunate to have had the opportunity to cross paths in life with every single one of you.
You are an incredibly bright and enjoyable group, and I am fortunate to have had the opportunity to cross paths in life with every single one of you.

>>>>

I have received an incredible amount of support during my time as an undergraduate student. 
First and foremost, I'd like to thank my advisor, Dr. Sergey Fomel, for his helpful guidance over the past couple of years. 
I genuinely believe that working with him has beneficially shaped the trajectory of my life.
Through his guidance and the opportunities he has given me, he has helped me to realize how much I truly enjoy learning, showed me the importance of reproducible computational research, introduced me to new ways of thinking about and approaching problems, and inspired me to continue my studies in graduate school.
For all that, I will be forever grateful.

@@

I would especially like to thank the members of the Texas Consortium for Computational Seismology (TCCS), of which I have been a part for the past two and a half years.
Thank you all for your intellectually stimulating conversations, life advice, and for including me as ``part of the group'', despite the fact that I was only an undergraduate student.
I will forever cherish the friendships and memories I have made with all of you during my time as a part of TCCS. 
You are an incredibly bright and enjoyable group, and I am fortunate to have had the opportunity to cross paths in life with every single one of you.

%I am very fortunate to be a part of the Jackson School of Geoscience's Undergraduate Honors Research Program, which has given me the opportunity to write this thesis. ... hm.

### 2018-05-25 chapter-background/paper.tex
thesis done

Chapter 3 is dedicated to the discussion of how to find this frequency balancing operator.
Figure \ref{fig:one1} shows the two traces after balancing the local frequency content of the red trace to match the black trace.

@@

\section{Shifting}
If two events are slightly misaligned in time, finding the transformation that correctly aligns these events is referred to as {\em shifting}.
Dynamic time warping is a common algorithm used for the alignment of events of two time series, both for geophysical and other applications, as diverse as speech recognition and finance \cite[]{herrera2012,hale2013,dtw,finance}.

In this thesis, I calculate the shifting operator by finding the time shifts that maximize local similarity between the two data sets.
Local similarity extends the concept of local seismic attributes to global correlation coefficients and is effective for measuring the match of seismic events in a local region of samples \cite[]{attr}.
Unlike filtering and scaling, shifting is a non-linear transformation.
%The time shift calculated is

@@

The result after finding and applying the time shift to the red trace to match it to the black trace is in Figure \ref{fig:one3}.
This shifting operation is discussed in more detail in chapters 3 and 4.

@@

The order and amount of these operators affects the final result.
After correcting for frequency variations, amplitude variations, and time shifts, additional corrections can be applied to further refine the match if necessary.
In some cases, multiple rounds of applying these operators and in different orders may be beneficial for the best match.
These three operators are noncommutative, so the order in which they are applied matters.
Applying filtering before scaling produces a different result from applying scaling before filtering. 
An illustration of this property is shown in Figure \ref{fig:bef,aft}.

@@

A summary of the three data matching operations is shown in Table \ref{op}.
Both scaling and filtering are linear operations, whereas shifting is generally a non-linear operation.
Additionally, as scaling is to weighting in the time domain, filtering is to weighting in the Fourier domain.
The next chapter discusses a few methods and examples of filtering, or frequency balancing, between two data sets.

>>>>

Chapter \ref{ch:chapter-locfreq} is dedicated to the discussion of how to find this frequency balancing operator.
Figure \ref{fig:one1} shows the two traces after balancing the local frequency content of the red trace to match the black trace.

@@

\section{Shifting}
If two events are misaligned in time, finding the transformation that correctly aligns these events is referred to as {\em shifting}.
Dynamic time warping is a common algorithm used for the alignment of events of two time series, both for geophysical and other applications, as diverse as speech recognition and finance \cite[]{herrera2012,hale2013,dtw,finance}.

In this thesis, I calculate the shifting operator by finding the time shifts that maximize local similarity between the two data sets.
Local similarity extends the concept of local seismic attributes to global correlation coefficients and is effective for measuring the match of seismic events in a local region of samples \cite[]{attr}.
%Unlike filtering and scaling, shifting is a non-linear transformation.
%The time shift calculated is

@@

The result after finding and applying the time shift to the red trace to match it to the black trace is in Figure \ref{fig:one3}.
This shifting operation is discussed in more detail in Chapters \ref{ch:chapter-locfreq} and \ref{ch:chapter-merge}..

@@

The order and amount of these operators affects the final result.
After correcting for frequency variations, amplitude variations, and time shifts, additional corrections can be applied to further refine the match if necessary.
These three operators are noncommutative, so the order in which they are applied matters.
Applying filtering before scaling produces a different result from applying scaling before filtering. 
An illustration of this property is shown in Figure \ref{fig:bef,aft}.
In some cases, multiple rounds of applying these operators and in different orders may be beneficial for the best match.

@@

A summary of the three data matching operations is shown in Table \ref{op}.
%Both scaling and filtering are linear operations, whereas shifting is generally a non-linear operation.
In the stationary case, if scaling is weighting in the time domain, filtering is weighting in the Fourier domain.
Chapter \ref{ch:chapter-locfreq} discusses a few methods and examples of filtering, or frequency balancing, between two data sets.

### 2018-05-25 chapter-locfreq/paper.tex
thesis done

        \subsection{Iterative method for calculating the smoothing radius}
        \inputdir{merge}
        The theoretical smoothing radius works in some situations, like in the first example of Chapter \ref{ch:chapter-merge}, but it does not work in situations where litte overlap is present in frequency content between data sets, like in the second example of Chapter \ref{ch:chapter-merge}.
        In this case, we use an iterative method to find what the smoothing radius should be that is based on the physical understanding of what smoothing is doing.

@@

        Using the assumptions, we can choose an intial guess for the smoothing radius and continually adjust the smoothing radius until we  achieve the desired result of balancing the local frequency content between the two data sets. 
        In practice, this method produces an acceptable solution in approximately 5 iterations and exhibits sublinear convergence. After smoothing the higher frequency data with the estimated radius, we use the lower frequency and smoothed higher frequency data to estimate time shifts and align the two data sets.

>>>>

        \subsection{Iterative method for calculating the smoothing radius}
        \inputdir{merge}
        The theoretical smoothing radius works in some situations, like in the first example of Chapter \ref{ch:chapter-merge}, but it does not work in situations where little overlap is present in frequency content between data sets, like in the second example of Chapter \ref{ch:chapter-merge}.
        In this case, we use an iterative method to find what the smoothing radius should be that is based on the physical understanding of what smoothing is doing.

@@

        Using the assumptions, we can choose an initial guess for the smoothing radius and continually adjust the smoothing radius until we  achieve the desired result of balancing the local frequency content between the two data sets. 
        In practice, this method produces an acceptable solution in approximately 5 iterations and exhibits sublinear convergence. After smoothing the higher frequency data with the estimated radius, we use the lower frequency and smoothed higher frequency data to estimate time shifts and align the two data sets.

### 2018-05-25 vita.tex
thesis done

Sarah graduated with high honors from Cinco Ranch High School in Katy, Texas in 2014 and came to The University of Texas at Austin (UT Austin) to pursue her undergraduate studies. She is a candidate for a Bachelor of Science degree in Geophysics from the Jackson School of Geosciences and a Bachelor of Science degree in Mathematics from the College of Natural Sciences. After graduation, she plans on pursuing a PhD in Mathematics and Computational Science at the Massachusetts Institute of Technology.

>>>>

Sarah graduated with high honors from Cinco Ranch High School in Katy, Texas in 2014 and came to The University of Texas at Austin to pursue her undergraduate studies. 
She is a candidate for a Bachelor of Science degree in Geophysics from the Jackson School of Geosciences and a Bachelor of Science degree in Mathematics from the College of Natural Sciences. 
After graduation, she plans on pursuing a PhD in Mathematics and Computational Science at the Massachusetts Institute of Technology.
%After graduation, she plans on pursuing a PhD in Computational Science, Engineering, and Mathematics at The University of Texas at Austin.
%After graduation, she plans on pursuing a PhD in Computational and Mathematical Engineering at Stanford University.
